\speculake: Hiding Malware in Speculative Execution

### 2018-08-07 paper/figures.tex
captions pass

\begin{figure}[ht]
    \centering
    % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
    \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-trimmed.pdf}
    \caption{\textbf{\speculake}\,---\,%
    Both the Trigger and Payload binaries exhibit the same initial series of
    indirect jumps, shown in Step 1. The trigger program solely executes these
    jumps to (mis)train the branch predictor. In the Payload program
    \texttt{fn\_ptr} has been set to point to the \texttt{cache\_timing}
    function and flushed from cache. Thus at Step 2 the branch predictor
    mis-predicts the jump to \texttt{cache\_timing} and instead jumps to
    \texttt{target\_fn} as trained by the Trigger program (green line). 
    \texttt{target\_fn} then executes speculatively as shown in Step 3. 
    Eventually \texttt{fn\_ptr} is loaded from memory and this mis-jump is 
    recognized by the processor, directing computation to Step 4, the 
    \texttt{cache\_timing} function, which then computes which value was loaded 
    into cache speculatively.}

    \label{fig:high-level}

@@

@@

\newcommand{\FigCacheMiss}{
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
    \caption{The Cache Hit/Miss Latency Distribution of an Intel
        Xeon-1270. Measured by timing accesses to cached/uncached memory. 
        Typical cache miss resolution takes 6x longer than a cache hit.}
    \label{fig:cache-miss}
\end{figure}
}

\newcommand{\FigGeneralModel}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
    \caption{General model of speculative computation within the Payload 
        process when triggered. \textit{Speculative Computation} has read-only
        access to all memory within the process, and returns a result via a
        cache side channel (via the \texttt{probe\_array}).
        The process can subsequently make
        \textit{Resultant Computations} based on the value returned from
        the cache side channel to update the state of the process. }
    \label{fig:general_model}
\end{figure}
}

\newcommand{\FigSpecBandwidth}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
    \caption{Speculative Bandwidth---Using the speculative primitive 1KB of data
        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
        the speculative world with 20 iterations per round for guaranteed
        correctness.  Varying the width of the channel results in competing
        performance factors -- decreasing the number of loops vs decreased size
        of the probe space.}
    \label{fig:spec_bandwidth}
\end{figure}
}

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
    \caption{This diagram demonstrates the control flow of a \speculake
        application making use of the decryption gadget and emulation. An
        encrypted SPASM binary stored in dead code or data accessible to
        \textit{Speculative Computation} is decrypted  and then returned through
        the primitive. The \textit{Resultant Computation} performs the SPASM
        emulation to update the process state before the next instruction is
        decrypted.}
    \label{fig:spasm_model}
\end{figure}
}

>>>>

\begin{figure}[ht]
    \centering
    % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
    \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-trimmed.pdf}
    \caption{\textbf{\speculake}\,---\,%
    Both the trigger and payload binaries exhibit the same initial series of
    indirect jumps (Step 1). The trigger program solely executes these
    jumps to (mis)train the branch predictor. In the Payload program
    \texttt{fn\_ptr} has been set to point to the \texttt{cache\_timing}
    function and flushed from cache. Thus the branch predictor
    mis-predicts the jump to \texttt{cache\_timing} and instead jumps to
    \texttt{target\_fn} (Step 2) as trained by the trigger program (green line). 
    \texttt{target\_fn} then executes speculatively (Step 3), until
    \texttt{fn\_ptr} is loaded from memory and the processor redirects
    computation to the \texttt{cache\_timing} function (Step 4),
    which measures which value was speculatively loaded into cache.}
    \label{fig:high-level}

@@

\newcommand{\FigSpectreOne}{
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
    \caption{\textbf{Speculative buffer overflow warm-up}\,---\, %
    The direct branch predictor must
    be trained to expect that a branch will go a specific way before speculative
    buffer overflows can be used. We varied the number of times a branch was trained
    to be taken and observed the fraction of times we achieved a speculative buffer overflow
    execution immediately after (measured by observing if a speculatively-loaded value was present
    in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
    direction at least 200 times before it can start to be used in a speculative buffer overflow.}
    \label{fig:spectre-one}
\end{figure}
}

\newcommand{\FigSpecMeasureNP}{
\begin{figure*}[t]
     \centering
        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
        \caption{\textbf{Speculative limits}\,---\, %
    To find the number of instructions we can execute speculatively, we placed a speculative load
    after an increasing number of instructions, and measured the fraction of times the load
    was subsequently in cache. We find that we can execute up to 220 instructions speculatively,
    which corresponds with the size of the CPU's reorder buffer (ROB). However, more complex instructions
    have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
    comprises.}
    \label{fig:spec-capacity-np}
\end{figure*}
}

\newcommand{\FigSpecMeasureParity}{
\begin{figure}[b]
     \centering
        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
        \caption{\textbf{Hyperthreading}\,---\, %
        The trigger program influences the payload program via the indirect branch predictor
        when running on the same logical CPU core. Hyperthreads also share an indirect branch
        predictor despite running on different logical cores. We measured the impact of
        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
        need to wait for an OS context switch between processes, but the two processes directly
        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
        speculative execution window.}
    \label{fig:spec-capacity-parity}
\end{figure}
}

@@

\newcommand{\FigCacheMiss}{
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
    \caption{\textbf{Cache latency}\,---\,%
        Cumulative distribution function of the cache hit and miss latency
        for an Intel Xeon-1270 and AMD Epyc 7251. If a cache miss is used
        to force CPU speculation, the CPU must wait at least 300-800 cycles
        before the speculated branch can be resolved. However, we find the CPU
        can execute far fewer instructions speculatively, suggesting another
        limit is at play.}
    \label{fig:cache-miss}
\end{figure}
}

\newcommand{\FigGeneralModel}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
        \caption{\textbf{ExSpectre model}\,---\, %
        General model of speculative computation within the payload 
        process when triggered. The \textit{Speculative Gadget} has read-only
        access to all memory within the process, but can return updates/results via a
        cache side channel (via access to the \texttt{probe\_array}).
        The process can subsequently make
        \textit{Resultant Computations} based on the value returned from
        the cache side channel to update the state of the \textit{Real World} process.}
    \label{fig:general_model}
\end{figure}
}

\newcommand{\FigSpecBandwidth}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
        \caption{\textbf{Speculative Bandwidth}\,---\, %
        Using the speculative primitive, 1KB of data
        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
        the speculative world with 20 iterations per round for guaranteed
        correctness.  Varying the width of the channel results in competing
        performance factors, with an optimal throughput achieved at 8~bits.}
    \label{fig:spec_bandwidth}
\end{figure}
}

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
        \caption{\textbf{SPASM model}\,---\, %
        Our SPASM emulator speculatively decrypts intructions, and emulates
        them in the real world. The \textit{Speculative Computation} decrypts
        the encrypted SPASM binary, returning the result through the side channel
        to allow the \textit{Resultant Computation} to update the emulated state
        (or make system calls) on behalf of the speculative world.}
    \label{fig:spasm_model}
\end{figure}
}
